\Rpackage{MSnbase}: labelled MS2 data pre-processing, visualisation and quantification.

### 2012-08-16 inst/doc/MSnbase-demo/MSnbase-demo.Rnw
not tikz usage anymore
git-svn-id: https://hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/MSnbase@68501 bc3139a8-67e5-0310-9ffc-ced21a209358

%\VignetteIndexEntry{Base Functions and Classes for MS-based Proteomics}
%\VignetteKeywords{Mass Spectrometry, MS, MSMS, Proteomics, Infrastructure, Bioinformatics, quantitative }
%\VignettePackage{MSnbase-demo}
\documentclass{article}
\usepackage{tikz}
\usepackage[nogin]{Sweave}
\usepackage{hyperref}
\usepackage[authoryear,round]{natbib}

@@

\begin{figure}[!ht]
<<spectrum-plot, dev='tikz', fig.width=4, fig.height=4, fig.keep='high'>>=
plot(sp, reporters = iTRAQ4, full = TRUE) 
@ 
\caption{Raw MS2 spectrum with details about reporter ions.}
\label{fig:spectrum-plot}
\end{figure}

@@

\begin{figure}[p]
<<msnexp-plot, dev='tikz', fig.width=3.7, fig.height=4.5, fig.keep='last'>>=
plot(bsa, reporters = iTRAQ4, full = FALSE) + theme_gray(8)
@ 
\caption{Experiment-wide raw MS2 spectra. The y-axes of the individual 
  spectra are automatically rescaled to the same range. 
  See section \ref{sec:norm} to rescale peaks identically. }
\label{fig:msnexp-plot}
\end{figure}

@@

\begin{figure}[!ht]
<<spectrum-clean-plot, dev='tikz', fig.width=6, fig.height=3, echo=FALSE, fig.keep='high'>>=
p1 <- plot(itraqdata[["X55"]], full = TRUE, plot = FALSE) + theme_gray(5) 
p2 <- plot(experiment[["X55"]], full = TRUE, plot = FALSE) + theme_gray(5) 
grid.newpage()
pushViewport(viewport(layout = grid.layout(1, 2)))
vplayout <- function(x, y)
  viewport(layout.pos.row = x, layout.pos.col = y)
print(p1,vp=vplayout(1,1))
print(p2,vp=vplayout(1,2))
@ 
\caption{Same spectrum before (left) and after setting peaks <= 400 to 0.}
\label{fig:spectrum-clean-plot}
\end{figure}

@@

\begin{figure}[p]
<<preproc-plot, dev='tikz', fig.width=3, fig.height=6, echo=FALSE>>=
grid.newpage()
pushViewport(viewport(layout = grid.layout(3, 1)))
print(p1, vp=vplayout(1,1))
print(p2, vp=vplayout(2,1))
print(p3, vp=vplayout(3,1))
@ 
\caption{This figure illustrated the effect or the \Rfunction{removePeaks} 
  and \Rfunction{clean} methods. The left-most spectrum displays two peaks, 
  of max height 3 and 7 respectively. The middle spectrum shows the result 
  of calling \Rfunction{removePeaks} with argument \Robject{t=3}, which sets 
  all data points of the first peak, whose maximum height is smaller or equal 
  to \Robject{t} to 0. The second peak is unaffected. Calling \Rfunction{clean} 
  after \Rfunction{removePeaks} effectively deletes successive 0 intensities from 
  the spectrum, as shown on the right plot. }
\label{fig:preproc}
\end{figure}

@@

\begin{figure}[p]
<<quantitation-plot, dev='tikz',fig.width=5, fig.height=5, echo=FALSE>>=
grid.newpage()
pushViewport(viewport(layout = grid.layout(2, 2)))
vplayout <- function(x, y)
  viewport(layout.pos.row = x, layout.pos.col = y)
print(p + opts(title = "Quantitation using 'sum'") +
      geom_point(size = 3, alpha = I(1/3), colour = "red"),
      vp = vplayout(1, 1))
print(p + opts(title = "Quantitation using 'max'") +
      geom_point(aes(x = 114.1, y = 3),
                 alpha = I(1/18),
                 colour = "red", size = 3),
      vp = vplayout(1, 2))
print(p +
      opts(title = "Trapezoidation and strict=FALSE") +
      geom_polygon(alpha = I(1/5), fill = "red"),
      vp = vplayout(2, 1))
print(p +
      opts(title = "Trapezoidation and strict=TRUE") + 
      geom_polygon(aes(x = c(NA, 114.05, 114.05, 114.1, 114.15, 114.15, NA),
                       y=c(NA, 0, 1, 3, 1, 0, NA)), fill = "red", alpha = I(1/5)),
      vp = vplayout(2,2))

@@

\begin{figure}[p]
<<norm-plot, dev='tikz', fig.width=5, fig.height=7, echo=FALSE>>=
.plot <- function(x,ttl=NULL) {
  boxplot(exprs(x),
          main=ifelse(is.null(ttl),processingData(x)@processing[2],ttl),
          cex.main=.8,
          cex.lab=.5,
          cex.axis=.5,
          cex=.8)
  grid()
}
oldmar <- par()$mar
par(mfrow=c(3,2),mar=c(2.9,2.9,2.9,1))
.plot(qnt, ttl = "Non-normalised data")
.plot(qnt.max, ttl = "Maximum")
.plot(qnt.sum, ttl = "Sum")
.plot(qnt.quant, ttl = "Quantile")
.plot(qnt.qrob, ttl = "Robust quantile")
.plot(qnt.vsn, ttl = "vsn")
@ 
\caption{ Comparison of the normalisation \Robject{MSnSet} methods. 
  Note that vsn also glog-transforms the intensities. }
\label{fig:norm-plot}
\end{figure}

@@

\begin{figure}[!ht]
<<cv-plot, dev='tikz', fig.width=5, fig.height=4, echo=FALSE>>=
print(p)
@ 
\caption{CV versus signal intensity comparison for log2 and vsn
  transformed data. Lines indicate running CV medians.}
\label{fig:cv-plot}
\end{figure}

@@

\begin{figure}[p]
<<msnexp-norm-plot, dev='tikz', fig.width=5, fig.height=6, echo=FALSE>>=
print(p)
@ 
\caption{Experiment-wide normalised MS2 spectra. The y-axes of the individual 
  spectra is now rescaled between 0 and 1 (highest peak), as opposed to 
  figure \ref{fig:msnexp-plot}. }
\label{fig:msnexp-norm-plot}
\end{figure}

@@

\begin{figure}[!hbt]
<<incompdiss-plot, dev='tikz', fig.width=5, fig.height=2.5, echo=FALSE, warning=FALSE>>=
dfr <- melt(exprs(incompdiss))
colnames(dfr) <- c("feature", "reporters", "intensity")
dfr$reporters <- sub("iTRAQ5.", "", dfr$reporters)
repsum <- apply(exprs(incompdiss), 1, function(x) sum(x[1:4]))
dfr2 <- data.frame(iTRAQ1to4 = repsum, iTRAQ5 = exprs(incompdiss)[,5])
p1 <- ggplot(data = dfr, aes(x = reporters,y = log10(intensity))) +
  geom_boxplot() +
  theme_gray(6)
p2 <- ggplot(data = dfr2, aes(x = log10(iTRAQ1to4), y = log10(iTRAQ5))) +
  geom_point(alpha = I(1/2)) +
  geom_abline(intercept = 0, slope = 1, linetype = "dotted") +
  stat_smooth(method = lm, se = FALSE) +
  xlab(label = expression(log[10]~sum~114~to~117)) +
  ylab(label = expression(log[10]~145)) +
  theme_gray(6)
grid.newpage()
pushViewport(viewport(layout = grid.layout(1, 2)))
print(p1, vp = vplayout(1, 1))
print(p2, vp = vplayout(1, 2))
@ 
\caption{Boxplot and scatterplot comparing intensities of the 4 reporter ions 
  (or their sum, on the right) and the incomplete dissociation specific peak. }
\label{fig:incompdiss}
\end{figure}

>>>>

%\VignetteIndexEntry{Base Functions and Classes for MS-based Proteomics}
%\VignetteKeywords{Mass Spectrometry, MS, MSMS, Proteomics, Infrastructure, Bioinformatics, quantitative }
%\VignettePackage{MSnbase-demo}
\documentclass{article}
\usepackage[nogin]{Sweave}
\usepackage{hyperref}
\usepackage[authoryear,round]{natbib}

@@

\begin{figure}[!ht]
<<spectrum-plot, dev='pdf', fig.width=4, fig.height=4, fig.keep='high'>>=
plot(sp, reporters = iTRAQ4, full = TRUE) 
@ 
\caption{Raw MS2 spectrum with details about reporter ions.}
\label{fig:spectrum-plot}
\end{figure}

@@

\begin{figure}[p]
<<msnexp-plot, dev='pdf', fig.width=3.7, fig.height=4.5, fig.keep='last'>>=
plot(bsa, reporters = iTRAQ4, full = FALSE) + theme_gray(8)
@ 
\caption{Experiment-wide raw MS2 spectra. The y-axes of the individual 
  spectra are automatically rescaled to the same range. 
  See section \ref{sec:norm} to rescale peaks identically. }
\label{fig:msnexp-plot}
\end{figure}

@@

\begin{figure}[!ht]
<<spectrum-clean-plot, dev='pdf', fig.width=6, fig.height=3, echo=FALSE, fig.keep='high'>>=
p1 <- plot(itraqdata[["X55"]], full = TRUE, plot = FALSE) + theme_gray(5) 
p2 <- plot(experiment[["X55"]], full = TRUE, plot = FALSE) + theme_gray(5) 
grid.newpage()
pushViewport(viewport(layout = grid.layout(1, 2)))
vplayout <- function(x, y)
  viewport(layout.pos.row = x, layout.pos.col = y)
print(p1,vp=vplayout(1,1))
print(p2,vp=vplayout(1,2))
@ 
\caption{Same spectrum before (left) and after setting peaks <= 400 to 0.}
\label{fig:spectrum-clean-plot}
\end{figure}

@@

\begin{figure}[p]
<<preproc-plot, dev='pdf', fig.width=3, fig.height=6, echo=FALSE>>=
grid.newpage()
pushViewport(viewport(layout = grid.layout(3, 1)))
print(p1, vp=vplayout(1,1))
print(p2, vp=vplayout(2,1))
print(p3, vp=vplayout(3,1))
@ 
\caption{This figure illustrated the effect or the \Rfunction{removePeaks} 
  and \Rfunction{clean} methods. The left-most spectrum displays two peaks, 
  of max height 3 and 7 respectively. The middle spectrum shows the result 
  of calling \Rfunction{removePeaks} with argument \Robject{t=3}, which sets 
  all data points of the first peak, whose maximum height is smaller or equal 
  to \Robject{t} to 0. The second peak is unaffected. Calling \Rfunction{clean} 
  after \Rfunction{removePeaks} effectively deletes successive 0 intensities from 
  the spectrum, as shown on the right plot. }
\label{fig:preproc}
\end{figure}

@@

\begin{figure}[p]
<<quantitation-plot, dev='pdf',fig.width=5, fig.height=5, echo=FALSE>>=
grid.newpage()
pushViewport(viewport(layout = grid.layout(2, 2)))
vplayout <- function(x, y)
  viewport(layout.pos.row = x, layout.pos.col = y)
print(p + opts(title = "Quantitation using 'sum'") +
      geom_point(size = 3, alpha = I(1/3), colour = "red"),
      vp = vplayout(1, 1))
print(p + opts(title = "Quantitation using 'max'") +
      geom_point(aes(x = 114.1, y = 3),
                 alpha = I(1/18),
                 colour = "red", size = 3),
      vp = vplayout(1, 2))
print(p +
      opts(title = "Trapezoidation and strict=FALSE") +
      geom_polygon(alpha = I(1/5), fill = "red"),
      vp = vplayout(2, 1))
print(p +
      opts(title = "Trapezoidation and strict=TRUE") + 
      geom_polygon(aes(x = c(NA, 114.05, 114.05, 114.1, 114.15, 114.15, NA),
                       y=c(NA, 0, 1, 3, 1, 0, NA)), fill = "red", alpha = I(1/5)),
      vp = vplayout(2,2))

@@

\begin{figure}[p]
<<norm-plot, dev='pdf', fig.width=5, fig.height=7, echo=FALSE>>=
.plot <- function(x,ttl=NULL) {
  boxplot(exprs(x),
          main=ifelse(is.null(ttl),processingData(x)@processing[2],ttl),
          cex.main=.8,
          cex.lab=.5,
          cex.axis=.5,
          cex=.8)
  grid()
}
oldmar <- par()$mar
par(mfrow=c(3,2),mar=c(2.9,2.9,2.9,1))
.plot(qnt, ttl = "Non-normalised data")
.plot(qnt.max, ttl = "Maximum")
.plot(qnt.sum, ttl = "Sum")
.plot(qnt.quant, ttl = "Quantile")
.plot(qnt.qrob, ttl = "Robust quantile")
.plot(qnt.vsn, ttl = "vsn")
@ 
\caption{ Comparison of the normalisation \Robject{MSnSet} methods. 
  Note that vsn also glog-transforms the intensities. }
\label{fig:norm-plot}
\end{figure}

@@

\begin{figure}[!ht]
<<cv-plot, dev='pdf', fig.width=5, fig.height=4, echo=FALSE>>=
print(p)
@ 
\caption{CV versus signal intensity comparison for log2 and vsn
  transformed data. Lines indicate running CV medians.}
\label{fig:cv-plot}
\end{figure}

@@

\begin{figure}[p]
<<msnexp-norm-plot, dev='pdf', fig.width=5, fig.height=6, echo=FALSE>>=
print(p)
@ 
\caption{Experiment-wide normalised MS2 spectra. The y-axes of the individual 
  spectra is now rescaled between 0 and 1 (highest peak), as opposed to 
  figure \ref{fig:msnexp-plot}. }
\label{fig:msnexp-norm-plot}
\end{figure}

@@

\begin{figure}[!hbt]
<<incompdiss-plot, dev='pdf', fig.width=5, fig.height=2.5, echo=FALSE, warning=FALSE>>=
dfr <- melt(exprs(incompdiss))
colnames(dfr) <- c("feature", "reporters", "intensity")
dfr$reporters <- sub("iTRAQ5.", "", dfr$reporters)
repsum <- apply(exprs(incompdiss), 1, function(x) sum(x[1:4]))
dfr2 <- data.frame(iTRAQ1to4 = repsum, iTRAQ5 = exprs(incompdiss)[,5])
p1 <- ggplot(data = dfr, aes(x = reporters,y = log10(intensity))) +
  geom_boxplot() +
  theme_gray(6)
p2 <- ggplot(data = dfr2, aes(x = log10(iTRAQ1to4), y = log10(iTRAQ5))) +
  geom_point(alpha = I(1/2)) +
  geom_abline(intercept = 0, slope = 1, linetype = "dotted") +
  stat_smooth(method = lm, se = FALSE) +
  xlab(label = expression(log[10]~sum~114~to~117)) +
  ylab(label = expression(log[10]~145)) +
  theme_gray(6)
grid.newpage()
pushViewport(viewport(layout = grid.layout(1, 2)))
print(p1, vp = vplayout(1, 1))
print(p2, vp = vplayout(1, 2))
@ 
\caption{Boxplot and scatterplot comparing intensities of the 4 reporter ions 
  (or their sum, on the right) and the incomplete dissociation specific peak. }
\label{fig:incompdiss}
\end{figure}
